Gravitational-wave sensitivity curves

### 2014-07-07 Introduction.tex
redone derivation of equation 7, now very long and detailed

Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$. The sensitivity of a detector to these depends upon the relative orientations of the source and detector. The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) the signal $h(t)$,
\begin{equation}
s(t) = n(t)+h(t) \; .
\end{equation}
We will work with the Fourier transform of the signal, using the conventions that
\begin{eqnarray} 
\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
\end{eqnarray}

@@

When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the average of the square of the detector noise:
\begin{eqnarray}\label{eq:meansquare}
\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft \right)\exp\left(-2\pi \rmi f't \right)
\end{eqnarray}
Using the Fourier property ${\mathcal{F}}\left\{ n(t+\tau) \right\} = \tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors may be absorbed as a time-translation of the noise by an amount $t$. The noise is stationary and is invariant under this time translation. We are thus taking the time average...
\cplb{I give up here. This derivation makes no sense in it's current state.}
\begin{eqnarray}\label{eq:meansquare1}
\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
\end{eqnarray}
Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of power spectral density in (\ref{eq:psd}) may be used.
\begin{eqnarray}\label{eq:meansquare1}
%\left<n(t)^{2}\right>&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; {\cal{F}}\left\{n(\tau+t)\right\}\left[{\cal{F}}\left\{n(\tau+t)\right\}\right]^{*}\nonumber \\
\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; \tilde{n}(f)\tilde{n}^{*}(f') \nonumber \\
&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
%&= \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta (f-f')\nonumber\\
& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
\end{eqnarray}
\cplb{Try someting based on Wiener--Khinchin theorem?}

>>>>

Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$. The sensitivity of a detector to these depends upon the relative orientations of the source and detector. The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) the signal $h(t)$,
\begin{equation}
s(t) = n(t)+h(t) \; .
\end{equation}
We will work with the Fourier transform of the signal, using the conventions that
\begin{eqnarray} \label{eq:defFourier}
\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
\end{eqnarray}

@@

When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the \emph{time} average of the square of the detector noise:
\begin{eqnarray}\label{eq:meansquare}
\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t) ^{2} \nonumber \\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \nonumber \\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')e^{2\pi \rmi ft }e^{-2\pi \rmi f't } \nonumber\\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau) \right\}(f){\cal{F}}\left\{ n(\tau) \right\}(f')^{*}e^{2\pi \rmi ft }e^{-2\pi \rmi f't }
\end{eqnarray}
Where in the penultimate line we have used the definition of the Fourier transform, (\ref{eq:defFourier}), and in the last line we have just changed notation. A property of Fourier transforms is that in the time domain a translation by amount $t$ translates to a rotation in phase by $2\pi ft$ in the frequency domain; if ${\mathcal{F}}\left\{ n(\tau) \right\} = \tilde{n}(f)$ then ${\mathcal{F}}\left\{ n(\tau-t) \right\} = \tilde{n}(f)\exp (2\pi \rmi ft )$. Therefore the exponential factors in (\ref{eq:meansquare}) may be absorbed as a time-translation of the noise by an amount $t$
\begin{eqnarray}
\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \, .
\end{eqnarray}
Using the ergodic property of the system described above the time average (denoted by $\overline{(\ldots)}$) may now be changed into an ensemble average (denoted by $\left<\ldots\right>$).
The noise is stationary so it's expectation value is unchanged by the time translation performed above. Therefore using (\ref{eq:psd}) gives
\begin{eqnarray}\label{eq:meansquare1}
&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
\end{eqnarray}
